Signatures and Induction Principles for Higher Inductive-Inductive Types

### 2019-01-31 paper.tex
Stuff

\newcommand{\Elim}{\mathsf{Elim}}
\newcommand{\elim}{\mathsf{elim}}
\newcommand{\Rec}{\mathsf{Rec}}
\newcommand{\record}{\mathsf{record}}
\newcommand{\funext}{\mathsf{funext}} \newcommand{\Q}{\mathsf{Q}}
\newcommand{\T}{\mathsf{T}} \newcommand{\leaf}{\mathsf{leaf}}
\newcommand{\node}{\mathsf{node}} \newcommand{\perm}{\mathsf{perm}}
\newcommand{\coe}{\mathsf{coe}} \newcommand{\vz}{\mathsf{vz}}
\newcommand{\vs}{\mathsf{vs}} \newcommand{\untr}{\mathsf{untr}}
\newcommand{\from}{\mathsf{from}}
\newcommand{\fromeq}{{\mathsf{from}\hspace{-0.3em}\equiv}}
\newcommand{\fromsimeq}{{\mathsf{from}\hspace{-0.3em}\simeq}}
\newcommand{\Model}{\mathsf{Model}}
\newcommand{\DModel}{\mathsf{DModel}}
\newcommand{\module}{\mathsf{module}}
\newcommand{\open}{\mathsf{open}}
\renewcommand{\P}{\mathsf{P}} \newcommand{\Bool}{\mathsf{Bool}}
\newcommand{\true}{\mathsf{true}} \newcommand{\false}{\mathsf{false}}
\renewcommand{\not}{\mathsf{not}} \newcommand{\0}{\mathsf{0}}
\newcommand{\1}{\mathsf{1}} \renewcommand{\Pr}{\mathsf{Pr}}
\newcommand{\PrNat}{\mathsf{PrNat}} \newcommand{\J}{\mathsf{J}}
\newcommand{\wkV}{\mathsf{wkV}} \renewcommand{\r}[1]{{\P_{#1}}}
\newcommand{\stab}{\mathsf{stab}} \newcommand{\NTy}{\mathsf{NTy}}
\newcommand{\isDec}{\mathsf{isDec}} \newcommand{\dec}{\mathsf{dec}}
\newcommand{\yes}{\mathsf{yes}} \newcommand{\no}{\mathsf{no}}
\newcommand{\case}{\mathsf{case}} \newcommand{\inj}{\mathsf{inj}}
\newcommand{\K}{\mathsf{K}} \newcommand{\lb}{\langle}
\newcommand{\rb}{\rangle} \newcommand{\Decl}{\mathsf{Decl}}
\newcommand{\Core}{\mathsf{Core}} \newcommand{\IND}{\mathsf{ind}}
\newcommand{\Id}{\mathsf{Id}} \newcommand{\Base}{\mathsf{Base}}
\newcommand{\transport}{\mathsf{transport}}
\newcommand{\I}{\mathsf{I}} \newcommand{\E}{\mathsf{S}}
\newcommand{\transp}{\mathsf{transp}}
\newcommand{\Transp}{\mathsf{Transp}}
\newcommand{\W}{\mathsf{W}}
\newcommand{\Fin}{\mathsf{Fin}} \newcommand{\fzero}{\mathsf{fzero}}
\newcommand{\fsuc}{\mathsf{fsuc}}
\newcommand{\inv}{\mathsf{inv}}
\newcommand{\con}{\mathsf{con}}
\newcommand{\LEFT}{\mathsf{left}}
\newcommand{\RIGHT}{\mathsf{right}}
\newcommand{\seg}{\mathsf{seg}}
\newcommand{\Int}{\mathsf{Int}}
\renewcommand{\in}{\mathbin{\hat:}}
\renewcommand{\hat}[1]{{\color{BrickRed}{#1}}}
\newcommand{\blc}[1]{{\color{Black}{#1}}}
\newcommand{\vdashh}{\mathbin{\hat\vdash}}
\newcommand{\rah}{\mathbin{\hat\ra}}
\newcommand{\commah}{\hat,\,}
\newcommand{\timesh}{\mathbin{\hat\times}}
\newcommand{\eqh}{\mathbin{\hat=}}
\newcommand{\TR}{\hat{\mathsf{tr}}}
\newcommand{\ap}{\hat{\mathsf{ap}}}
\newcommand{\apd}{\hat{\mathsf{apd}}}
\renewcommand{\tt}{\hat{\mathsf{tt}}}
\newcommand{\Tel}{\hat{\mathsf{Tel}}}
\newcommand{\Type}{\hat{\mathsf{Type}}}
\newcommand{\emptytel}{\hat{\epsilon}}
\newcommand{\telext}{\mathbin{\hat{\lhd}}}
\newcommand{\ltel}{\hat{\ll}}
\newcommand{\rtel}{\hat{\rr}}
\newcommand\pp{\ensuremath{\hat{\mathbin{+\mkern-10mu+}}}}
\newcommand{\semicol}{\hat;\,}
\newcommand{\targetass}{\hat{\Gamma}\semicol}
\newcommand{\Alg}{\mathsf{Alg}}
\newcommand{\DisplayedAlg}{\mathsf{DisplayedAlg}}
\newcommand{\Morphism}{\mathsf{Morphism}}
\newcommand{\Section}{\mathsf{Section}}
\newcommand{\ass}{\mathsf{ass}}
\newcommand{\idl}{\mathsf{idl}}
\newcommand{\idr}{\mathsf{idr}}
\newcommand{\inv}{\mathsf{inv}}

@@

We can easily show that the non-truncated syntax does not form sets. For
example, $\U[]$ and $[\id]$ are two proofs of $\U[\id] = \U$, and they are not
forced to be equal. Assuming univalence, we can give a model where types are
interpreted as closed metatheoretic types, $\U$ is interpreted as metatheoretic
$\Bool$, $[\id]$ is interpreted as $\refl : A = A$ for some metatheoretic $A$
type, and $\U[]$ is interpreted as the $\Bool$ negation equivalence, thereby
formally distinguishing $\U[]$ and $[\id]$. Hence, by Hedberg's theorem
\cite{hedberg}, the syntax does not have decidable equality. This also implies
that the non-truncated intrinsic syntax is not constructible from set quotients
of extrinsic terms (since those always form sets). The non-truncated syntax just
does not seem to be a sensible notion. This situation is similar to how
categories in homotopy type theory need to have set-truncated objects
\cite{ahrens2015univalent}.

@@

In the following three sections we present syntactic translations yielding
algebras, homomorphisms, displayed algebras and their sections. The presentation
here is informal and focuses on readability, similarly to our presentation of
signatures before.

@@

\begingroup
\allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\A && :\equiv \hat{\top} \\
  & (\Delta,x:A)^\A && :\equiv \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\timesh A^\A\,\hat{\gamma} \\
  & x^\A\,\hat{\gamma} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma} \\ % TODO: write De Bruijn versions
  & \U^\A\,\hat{\gamma} && :\equiv \Type_{\hat{0}} \\
  & (\underline{a})^\A\,\hat{\gamma} && :\equiv a^\A\,\hat{\gamma} \\
  & ((x:a)\ra B)^\A\,\hat{\gamma} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\rah  B^\A\,\hat{(}\hat{\gamma}\commah\hat{x}\hat{)} \\
  & (t\,u)^\A\,\hat{\gamma} && :\equiv \hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{(}u^\A\,\hat{\gamma}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\A\,\hat{\gamma} && :\equiv \hat{(}\hat{x}\in \hat{A})\rah  B^\A\,\hat{\gamma} \\
  & (t\,\hat{u})^\A\,\hat{\gamma} && :\equiv \hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{u} \\
  & (t=_a u)^\A\,\hat{\gamma} && :\equiv t^\A\,\hat{\gamma} \eqh u^\A\,\hat{\gamma} \\
  & \refl^\A\,\hat{\gamma} && :\equiv \hat{\refl} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\A\,\hat{\gamma} && :\equiv \hat{\J}_{\hat{(}a^\A\,\hat{\gamma}\hat{)}\,\hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{(}\hat{\lambda} \hat{x}\,\hat{z}.p^\A\,\hat{(}\hat{\gamma}\commah\hat{x}\commah\hat{z}\hat{)}\hat{)}}\,\hat{(}pr^\A\,\hat{\gamma}\hat{)}\,_{\hat{(}u^\A\,\hat{\gamma}\hat{)}}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\A\,\hat{\gamma} && :\equiv \hat{\refl} \\
  & ((\hat{x}\in \hat{A})\ra b)^\A\,\hat{\gamma} && :\equiv (\hat{x}\in \hat{A}\hat{)}\rah  b^\A\,\hat{\gamma} \\
  & (t\,\hat{u})^\A\,\hat{\gamma} && :\equiv \hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{u}
\end{alignat*}
\endgroup

@@

In this section, we specify the $\blank^\M$ translation, which computes
homomorphisms of algebras. We do so by first considering a logical relation
interpretation, and then refining it towards computing homomorphisms. For
dependent type theories, logical relation models are well-known (see
e.g.\ \cite{atkey}), so they are certainly applicable to our restricted syntax
as well.

Below, we list induction motives for $\blank^\M$; these remain the same as we
move from logical relations to homomorphisms. We fix a universe level $\hat{i}$ for
the translations.
\[
\infer{\hat{\Gamma}\vdashh\Delta^\M \in \Delta^\A\rah\Delta^\A\rah \Type_{\hat{i}}}
      {\hat{\Gamma}\vdash\Delta}
\hspace{1em}
\infer{
  \hat{\Gamma}\vdashh A^\D \in \hat{(}\hat{\gamma_0\,\gamma_1}\in\Delta^\A\hat{)}\rah \Delta^\M\,\hat{\gamma_0\,\gamma_1}\rah  A^\A\,\hat{\gamma_0}\rah A^\A\,\hat{\gamma_1}\rah \Type_{\hat{i}}}
      {\hat{\Gamma}\semicol\Delta\vdash A}
\]

\[
\infer{
  \hat{\Gamma}\vdashh A^\D \hat{ : (\gamma_0\,\gamma_1 : \blc{\Delta^\A})(\gamma^M : \blc{\Delta^\M}\,\gamma_0\,\gamma_1)\ra \blc{A^\M}\,\gamma_0\,\gamma_1\,\gamma^M\,(\blc{t^\A}\,\gamma_0)\,(\blc{t^\A}\,\gamma_1)
      }}
      {\hat{\Gamma}\semicol\Delta\vdash t : A}
\]

Contexts are mapped to (proof-relevant) relations and types to families of relations
depending on interpreted contexts. For terms, we get something which is often called
a \emph{fundamental theorem} of a logical relation: every term is related to itself
in related semantic contexts.

We interpret the universe as relation space, and function types as types
expressing pointwise relatedness of functions. This interpretation works the
same way for unrestricted (non-strictly positive) function types as well.

However, we would like to have underlying functions instead of relations in
homomorphisms. We take hint from the fact that for classical (simply-typed)
algebraic theories, a logical relation is equivalent to a homomorphism if and
only if the underlying relation is the graph of a function
\cite[pg. 5]{udayReynolds}. Thus we make the following change:
\begin{alignat*}{5}
  & \U^\M\,\hat{\gamma_0\,\gamma_1\,\gamma^M\,A_0\,A_1} && :\equiv \hat{A_0\ra A_1}
\end{alignat*}
This requires us to change $(\underline{a})^\M$ as well, since we need to
produce a type as result, but $a^M$ now yields a function. We can morally keep
the old definition, by viewing the result of $a^M$ as a functional relation, and
relating $t_0$ and $t_1$ by its graph:
\begin{alignat*}{5}
  & (\underline{a})^\M\,\hat{\gamma_0\,\gamma_1\,\gamma^M\,t_0\,\,t_1} && :\equiv \hat{
    \blc{a^\M}\,\gamma_0\,\gamma_1\,\gamma^M\,t_0 = t_1}
\end{alignat*}
At this point we have merely restricted relations to functions, and left the
rest of the interpretation unchanged. However, this is not strictly the desired
notion of homomorphism. Consider the $\blank^\M$ interpretation for natural
numbers:
\begin{alignat*}{5}
  & && (\boldsymbol{\cdot},Nat : \U,zero:\underline{Nat},suc:Nat\ra\underline{Nat})^\M
    \,\hat{(\tt,\,N_0,\,z_0,\,s_0)\,(\tt,\,N_1,\,z_1,\,s_1)} \\
    & \equiv\,\, && \hat{\top\times(N^M : N_0\ra N_1)}\\
    & && \hat{\hspace{1em}\times\,\,(z^M : N^M\,z_0 = z_1)}\\
    & && \hat{\hspace{1em}\times\,\,(s^M : (x_0 : N_0)(x_1 : N_1)(x^M : N^M\,x_0 = x_1)\ra N^M\,(s_0\,x_0)=s_1\,x_1)}
\end{alignat*}
In $\hat{s^M}$, there is a superfluous $\hat{x^M}$ equality proof. Fortunately,
in the translation of the inductive function space, we can just singleton
contract $\hat{x^M}$ away, yielding an equivalent, but stricter definition:
\begin{alignat*}{5}
  & ((x:a)\ra B)^\M\,\hat{\gamma_0\,\gamma_1\,\gamma^M\,f_0\,f_1} :\equiv \\
  & \hat{\hspace{2em}(x_0 : \blc{a^\A}\,\gamma_0)\ra \blc{B^\M}\,(\gamma_0,\,x_0)\,(\gamma_1,\,\blc{a^\M}\,\gamma_0\,\gamma_1\,\gamma^M\,x_0)\,(\gamma^M,\,\refl)\,
    (f_0\,x_0)\,(f_1\,(\blc{a^\M}\,\gamma_0\,\gamma_1\,\gamma^M\,x_0))}
\end{alignat*}
Now, the $\beta$-rule for successors is as expected:
\begin{alignat*}{5}
 & \hat{s^M : (x_0 : N_0)\ra N^M\,(s_0\,x_0)=s_1\,(N^M\,x_0)}
\end{alignat*}

$\blank^\M$ is very similar to the translation to displayed algebra sections,
which is discussed in Section \ref{sec:sections}. In essence, $\blank^\M$ is a
non-dependently typed version of the section translation. We will discuss in
more detail the interpretations of equalities and the other function types in
Section \ref{sec:sections}, and we leave the rest of the homomorphism
translation to Appendix \ref{sec:morphismrules}.

@@

The $\blank^\D$ translation computes displayed algebras for a signature, which
can be viewed as bundles of induction motives and methods. $\blank^\D$ is
essentially an unary logical predicate translation over the $\blank^\A$
model. It is related to the logical predicate translation of Bernardy et
al.\ \cite{bernardy12parametricity}, but our implementation differs by
interpreting contexts as $\Sigma$-types instead of extended contexts.

For each
context $\Delta$, $\Delta^\D$ is a predicate over the standard interpretation
$\Delta^\A$. For a type $\Delta\vdash A$, $A^\D$ is a predicate over $A^\A$,
which also depends on $\hat{\gamma}\in\Delta^\A$ and a witness of
$\Delta^\D\,\hat{\gamma}$. All of these may refer to a target theory context
$\hat{\Gamma}$.
\[
\infer{\hat{\Gamma}\vdashh\Delta^\D \in \Delta^\A\rah \Type_{\hat{i+1}}}{\hat{\Gamma}\vdash\Delta}
\hspace{2em}
\infer{
  \hat{\Gamma}\vdashh A^\D \in \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\rah \Delta^\D\,\hat{\gamma}\rah  A^\A\,\hat{\gamma}\rah \Type_{\hat{i+1}}}
      {\hat{\Gamma}\semicol\Delta\vdash A}
\]
For a term $t$, $t^\D$ witnesses that the predicate corresponding to its type
holds for $t^\A$; again this can be viewed as a fundamental theorem for the
predicate interpretation.

>>>>

\newcommand{\Elim}{\mathsf{Elim}}
\newcommand{\elim}{\mathsf{elim}}
\newcommand{\Rec}{\mathsf{Rec}}
\newcommand{\record}{\mathsf{record}}
\newcommand{\funext}{\mathsf{funext}} \newcommand{\Q}{\mathsf{Q}}
\newcommand{\T}{\mathsf{T}} \newcommand{\leaf}{\mathsf{leaf}}
\newcommand{\node}{\mathsf{node}} \newcommand{\perm}{\mathsf{perm}}
\newcommand{\coe}{\mathsf{coe}} \newcommand{\vz}{\mathsf{vz}}
\newcommand{\vs}{\mathsf{vs}} \newcommand{\untr}{\mathsf{untr}}
\newcommand{\from}{\mathsf{from}}
\newcommand{\fromeq}{{\mathsf{from}\hspace{-0.3em}\equiv}}
\newcommand{\fromsimeq}{{\mathsf{from}\hspace{-0.3em}\simeq}}
\newcommand{\Model}{\mathsf{Model}}
\newcommand{\DModel}{\mathsf{DModel}}
\newcommand{\module}{\mathsf{module}}
\newcommand{\open}{\mathsf{open}}
\renewcommand{\P}{\mathsf{P}} \newcommand{\Bool}{\mathsf{Bool}}
\newcommand{\true}{\mathsf{true}} \newcommand{\false}{\mathsf{false}}
\renewcommand{\not}{\mathsf{not}} \newcommand{\0}{\mathsf{0}}
\newcommand{\1}{\mathsf{1}} \renewcommand{\Pr}{\mathsf{Pr}}
\newcommand{\PrNat}{\mathsf{PrNat}} \newcommand{\J}{\mathsf{J}}
\newcommand{\wkV}{\mathsf{wkV}} \renewcommand{\r}[1]{{\P_{#1}}}
\newcommand{\stab}{\mathsf{stab}} \newcommand{\NTy}{\mathsf{NTy}}
\newcommand{\isDec}{\mathsf{isDec}} \newcommand{\dec}{\mathsf{dec}}
\newcommand{\yes}{\mathsf{yes}} \newcommand{\no}{\mathsf{no}}
\newcommand{\case}{\mathsf{case}} \newcommand{\inj}{\mathsf{inj}}
\newcommand{\K}{\mathsf{K}} \newcommand{\lb}{\langle}
\newcommand{\rb}{\rangle} \newcommand{\Decl}{\mathsf{Decl}}
\newcommand{\Core}{\mathsf{Core}} \newcommand{\IND}{\mathsf{ind}}
\newcommand{\Id}{\mathsf{Id}} \newcommand{\Base}{\mathsf{Base}}
\newcommand{\transport}{\mathsf{transport}}
\newcommand{\I}{\mathsf{I}} \newcommand{\E}{\mathsf{S}}
\newcommand{\transp}{\mathsf{transp}}
\newcommand{\Transp}{\mathsf{Transp}}
\newcommand{\W}{\mathsf{W}}
\newcommand{\Fin}{\mathsf{Fin}} \newcommand{\fzero}{\mathsf{fzero}}
\newcommand{\fsuc}{\mathsf{fsuc}}
\newcommand{\inv}{\mathsf{inv}}
\newcommand{\con}{\mathsf{con}}
\newcommand{\LEFT}{\mathsf{left}}
\newcommand{\RIGHT}{\mathsf{right}}
\newcommand{\seg}{\mathsf{seg}}
\newcommand{\Int}{\mathsf{Int}}
\renewcommand{\in}{\mathbin{\hat:}}
\renewcommand{\hat}[1]{{\color{BrickRed}{#1}}}
\newcommand{\blc}[1]{{\color{Black}{#1}}}
\newcommand{\vdashh}{\mathbin{\hat\vdash}}
\newcommand{\rah}{\mathbin{\hat\ra}}
\newcommand{\commah}{\hat,\,}
\newcommand{\timesh}{\mathbin{\hat\times}}
\newcommand{\eqh}{\mathbin{\hat=}}
\newcommand{\TR}{\hat{\mathsf{tr}}}
\newcommand{\ap}{\hat{\mathsf{ap}}}
\newcommand{\apd}{\hat{\mathsf{apd}}}
\renewcommand{\tt}{\hat{\mathsf{tt}}}
\newcommand{\Tel}{\hat{\mathsf{Tel}}}
\newcommand{\Type}{\hat{\mathsf{Type}}}
\newcommand{\emptytel}{\hat{\epsilon}}
\newcommand{\telext}{\mathbin{\hat{\lhd}}}
\newcommand{\ltel}{\hat{\ll}}
\newcommand{\rtel}{\hat{\rr}}
\newcommand\pp{\ensuremath{\hat{\mathbin{+\mkern-10mu+}}}}
\newcommand{\semicol}{\hat;\,}
\newcommand{\targetass}{\hat{\Gamma}\semicol}
\newcommand{\Alg}{\mathsf{Alg}}
\newcommand{\DisplayedAlg}{\mathsf{DisplayedAlg}}
\newcommand{\Morphism}{\mathsf{Morphism}}
\newcommand{\Section}{\mathsf{Section}}
\newcommand{\ass}{\mathsf{ass}}
\newcommand{\idl}{\mathsf{idl}}
\newcommand{\idr}{\mathsf{idr}}
\newcommand{\inv}{\mathsf{inv}}
\newcommand{\ID}{\mathsf{\ID}}

@@

We can easily show that the non-truncated syntax does not form sets. For
example, $\U[]$ and $[\id]$ are two proofs of $\U[\id] = \U$, and they are not
forced to be equal. Assuming univalence, we can give a model where types are
interpreted as closed metatheoretic types, $\U$ is interpreted as metatheoretic
$\Bool$, $[\id]$ is interpreted as $\refl : A = A$ for some metatheoretic $A$
type, and $\U[]$ is interpreted as the $\Bool$ negation equivalence, thereby
formally distinguishing $\U[]$ and $[\id]$. Hence, by Hedberg's theorem
\cite{hedberg}, the syntax does not have decidable equality. This also implies
that the non-truncated intrinsic syntax is not constructible from set quotients
of extrinsic terms (since those always form sets). The non-truncated syntax just
does not seem to be a sensible notion. This situation is similar to how
categories in homotopy type theory need to have set-truncated morphisms
\cite{ahrens2015univalent}.

@@

In the following three sections we present syntactic translations yielding
algebras, homomorphisms, displayed algebras and their sections. The presentation
here, like in Section \ref{sec:tos}, is informal and focuses on readability.
In particular, we omit interpretations for substitutions and preservation
proofs for definitional equalities.

@@

\begingroup
\allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\A && :\equiv \hat{\top} \\
  & (\Delta,x:A)^\A && :\equiv \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\timesh A^\A\,\hat{\gamma} \\
  & x^\A\,\hat{\gamma} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma} \\ % TODO: write De Bruijn versions
  & \U^\A\,\hat{\gamma} && :\equiv \Type_{\hat{0}} \\
  & (\underline{a})^\A\,\hat{\gamma} && :\equiv a^\A\,\hat{\gamma} \\
  & ((x:a)\ra B)^\A\,\hat{\gamma} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\rah  B^\A\,\hat{(}\hat{\gamma}\commah\hat{x}\hat{)} \\
  & (t\,u)^\A\,\hat{\gamma} && :\equiv \hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{(}u^\A\,\hat{\gamma}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\A\,\hat{\gamma} && :\equiv \hat{(}\hat{x}\in \hat{A})\rah  B^\A\,\hat{\gamma} \\
  & (t\,\hat{u})^\A\,\hat{\gamma} && :\equiv \hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{u} \\
  & (t=_a u)^\A\,\hat{\gamma} && :\equiv t^\A\,\hat{\gamma} \eqh u^\A\,\hat{\gamma} \\
  & \refl^\A\,\hat{\gamma} && :\equiv \hat{\refl} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\A\,\hat{\gamma} && :\equiv \hat{\J}_{\hat{(}a^\A\,\hat{\gamma}\hat{)}\,\hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{(}\hat{\lambda} \hat{x}\,\hat{z}.p^\A\,\hat{(}\hat{\gamma}\commah\hat{x}\commah\hat{z}\hat{)}\hat{)}}\,\hat{(}pr^\A\,\hat{\gamma}\hat{)}\,_{\hat{(}u^\A\,\hat{\gamma}\hat{)}}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\A\,\hat{\gamma} && :\equiv \hat{\refl} \\
  & ((\hat{x}\in \hat{A})\ra b)^\A\,\hat{\gamma} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  b^\A\,\hat{\gamma} \\
  & (t\,\hat{u})^\A\,\hat{\gamma} && :\equiv \hat{(}t^\A\,\hat{\gamma}\hat{)}\,\hat{u}
\end{alignat*}
\endgroup

@@

@@

The $\blank^\D$ translation computes displayed algebras for a signature, which
can be viewed as bundles of induction motives and methods. $\blank^\D$ is an
unary logical predicate translation over the $\blank^\A$ translation. It is
related to the logical predicate translation of Bernardy et
al.\ \cite{bernardy12parametricity}, but our implementation differs by
interpreting contexts as $\Sigma$-types instead of extended contexts.

We fix a universe level $\hat{i}$ for the translation. For each context
$\Delta$, $\Delta^\D$ is a predicate over the standard interpretation
$\Delta^\A$. For a type $\Delta\vdash A$, $A^\D$ is a predicate over $A^\A$,
which also depends on $\hat{\gamma}\in\Delta^\A$ and a witness of
$\Delta^\D\,\hat{\gamma}$. All of these may refer to a target theory context
$\hat{\Gamma}$.
\[
\infer{\hat{\Gamma}\vdashh\Delta^\D \in \Delta^\A\rah \Type_{\hat{i+1}}}{\hat{\Gamma}\vdash\Delta}
\hspace{2em}
\infer{
  \hat{\Gamma}\vdashh A^\D \in \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\rah \Delta^\D\,\hat{\gamma}\rah  A^\A\,\hat{\gamma}\rah \Type_{\hat{i+1}}}
      {\hat{\Gamma}\semicol\Delta\vdash A}
\]
For a term $t$, $t^\D$ witnesses that the predicate corresponding to its type
holds for $t^\A$; this can be viewed as a \emph{fundamental theorem} for the
predicate interpretation.

\[
\infer{\hat{\Gamma}\vdashh t^\D \in \hat{(}\hat{\gamma}\in\Delta^\A\hat{)}\hat{(}\hat{\gamma^D}\in\Delta^\D\,\hat{\gamma}\hat{)}\rah  A^\D\,\hat{\gamma}\,\hat{\gamma^D}\,\hat{(}t^\A\,\hat{\gamma}\hat{)}}{\hat{\Gamma}\semicol\Delta\vdash t : A}
\]
The implementation of $\blank^\D$ is given below. We leave
$\hat{\gamma}$-s mostly implicit, marking only $\hat{\gamma^D}$ witnesses.
\begingroup
%% \allowdisplaybreaks
\begin{alignat*}{5}
  & \boldsymbol{\cdot}^\D\,\hat{\gamma} && :\equiv \hat{\top} \\
  & (\Delta,\,x:A)^\D\,(\hat{\gamma}\commah\,\hat{t}) && :\equiv \hat{(}\hat{\gamma^D}\in\Delta^D\,\hat{\gamma}\hat{)}\timesh A^D\,\hat{\gamma^D}\,\hat{t} \\
  & x^\D\,\hat{\gamma^D} && :\equiv x^{\text{th}}\text{ component in } \hat{\gamma^D} \\
  & \U^\D\,\hat{\gamma^D}\,\hat{A} && :\equiv \hat{A} \rah  \Type_{\hat{i}} \\
  & (\underline{a})^\D\,\hat{\gamma^D}\,\hat{t} && :\equiv a^\D\,\hat{\gamma^D}\,\hat{t} \\
  & ((x:a)\ra B)^\D\,\hat{\gamma^D}\,\hat{f} && :\equiv \hat{(}\hat{x}\in a^\A\,\hat{\gamma}\hat{)}\hat{(}\hat{x^D}\in a^\D\,\hat{\gamma^D}\,\hat{x}\hat{)} \rah  B^\D\,\hat{(}\hat{\gamma}\commah\hat{x}\hat{)}\,\hat{(}\hat{\gamma^D}\commah\hat{x^D}\hat{)}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,u)^\D\,\hat{\gamma^D} && :\equiv \hat{(}t^\D\,\hat{\gamma^D}\hat{)}\,\hat{(} u^\A\,\hat{\gamma}\hat{)}\,\hat{(}u^\D\,\hat{\gamma^D}\hat{)} \\
  & ((\hat{x}\in \hat{A})\ra B)^\D\,\hat{\gamma^D}\,\hat{f} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  B^\D\,\hat{\gamma^D}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\D\,\hat{\gamma^D} && :\equiv t^\D\,\hat{\gamma^D}\,\hat{u} \\
  & (t=_a u)^\D\,\hat{\gamma^D}\,\hat{e} && :\equiv \TR_{\hat{(}a^\D\,\hat{\gamma^D}\hat{)}}\,\hat{e}\,\hat{(}t^\D\,\hat{\gamma^D}\hat{)} \eqh u^\D\hat{\gamma^D} \\
  & (\refl_t)^\D\,\hat{\gamma^D} && :\equiv \hat{\refl}_{\hat{(}t^\D\,\hat{\gamma^D}\hat{)}} \\
  & (\J_{a\,t\,(x.z.p)}\,pr\,_u\,eq)^\D\,\hat{\gamma^D} && :\equiv \hat{\J}\,\hat{\big(}\hat{\J}\,\hat{(}pr^\D\,\hat{\gamma^D}\hat{)}\,\hat{(}eq^\A\,\hat{\gamma}\hat{)}\hat{\big)}\,\hat{(}eq^\D\,\hat{\gamma^D}\hat{)} \\
  & (\J\beta_{a\,t\,(x.z.p)}\,pr)^\D\,\hat{\gamma^D} && :\equiv \hat{\refl} \\
  & ((\hat{x}\in \hat{A})\ra b)^\D\,\hat{\gamma^D}\,\hat{f} && :\equiv \hat{(}\hat{x}\in \hat{A}\hat{)}\rah  b^D\,\hat{\gamma^D}\,\hat{(}\hat{f}\,\hat{x}\hat{)} \\
  & (t\,\hat{u})^\D\,\hat{\gamma^D} && :\equiv t^\D\,\hat{\gamma^D}\,\hat{u}
\end{alignat*}
\endgroup
The predicate for a context is given by iterating $\blank^\D$ for its
constituent types. For a variable, the corresponding witness is looked up from
$\hat{\gamma^D}$.
